A Critical Review of Sophistication and Model Selection in Algorithmic Statistics

### 2014-12-09 Facticity/Facticity/paper.tex
...

\begin{definition}[Candidate set]
  The \emph{candidate models} for $x$ from model class $\M$ and
  threshold function $\tau$ are
  \[
  C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x)\}.
  \]
\end{definition}

@@

\begin{itemize}
\item Model class is $\C$
\item Description method for models is $L(m)=|\bar\imath|$ where
  $\phi_i=m$. The index function is assumed to be \emph{faithful}, see
  Definition~\ref{def:faithful}
\item Threshold function is $\tau(x)=L^m(x)$.
\item Facticity is actually defined as index length without prefix
  encoding:  $S^\C_0(x):=\min\{|i|:\phi_i\in C^\C_0(x)\}$. But this
  difference is not important for the discussion.
\end{itemize}

@@

@@

Proof sketch:

\begin{itemize}
  \item There is a $C$ such that for every $n$ there are strings $x$
    with $K(S)+\log|S|\ge K(x)+(3/4)n+C$ for all $K(S)>1/4$.

  \item For all $C$ there is an $n$ such that there are $x$ with
    $K(S)+\log(S)<K(x)+C$ only if $K(S)>1/4$.
\end{itemize}

>>>>

\begin{definition}[Candidate set]
  The \emph{candidate models} for $x$ from model class $\M$ and
  threshold function $\tau$ are
  \[
  C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x) + c\}
  \]
  with $c$ a constant independent of $x$ which may be chosen arbitrarily. The default threshold function is $\tau^\circ(x)=L^\M(x)$.
\end{definition}

@@

\begin{itemize}
\item Model class is $\C$
\item Description method for models is $L(m)=|\bar\imath|$ where
  $\phi_i=m$. The index function is assumed to be \emph{faithful}, see
  Definition~\ref{def:faithful}
\item Threshold function is $\tau(x)=L^\M(x)$, the constant in the candidate set is $0$.
\item Facticity is actually defined as index length without prefix
  encoding:  $S^\C_0(x):=\min\{|i|:\phi_i\in C^\C_0(x)\}$. But this
  difference is not important for the discussion.
\end{itemize}

@@

\begin{lemma}
Let $f$ be a total, computable function, such that for some $d$, $f(d) = x$ and let $k = K(f) + |d|$. Then there exists a finite set $S$ containing $x$ and a constant $c$, such that $K(S) \leq K(f) + K(|d|) + c$ and $\log |S| \leq |d|$.\label{lemma:total-to-sets}
\end{lemma}
\begin{proof}
Let $S = f\left(\{0,1\}^{|d|}\right)$. Since $f$ is total, this set can be explicity computed from a description of $f$ and a description of $|d|$, which tells us that there is a constant $c$ such that $K(S) \leq K(f) + K(|d|) + c$.

Since $S$ is the image of a set of size $2^{|d|}$ under $f$, we have $\log |S| \leq |d|$.
\end{proof}
This lemma is a variation on \cite[Lemma~7.2]{vitanyi2004meaningful}.

\begin{theorem}
There are infinitely many $x$ with 
\begin{enumerate}
  \item $S^\F_K(x) \geq K(x)/5$,\label{eq:item1}
  \item $S^\T_K(x) \geq K(x)/6$, {and}\label{eq:poezenvoer}
  \item $S^\T_{\tau^\circ}(x) \geq K(x)/6$. \label{eq:hondevoer}
\end{enumerate}
\end{theorem}
\begin{proof}
From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic. For each $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic. Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this sequence contains infinitely many distinct binary strings. Now pick any $i$ that is large enough that (a) $|x_i|<i$ and (b) $K(x_i)/5 < i/4$. By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $S$ containing $x$, either (a) $\log|S|\ge K(x_i)+i/4$, in which case $S$ cannot be a candidate representation for $x$ in $C^\F_K$, or (b) $K(S)>i/4>K(x_i)/5$. This proves Item~\ref{eq:item1}.

For Items~\ref{eq:poezenvoer} and~\ref{eq:hondevoer}, suppose towards contradiction that either $C^\T_K(x)$ or $C^T_{\tau^\circ}(x)$ contains a candidate $f$ with $K(f)\le i/5$. Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\le\tau^\circ(x)\le K(x)$, so (a) above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) for large enough $i$. Therefore such candidates $f$ cannot exist and all candidates must satisfy $K(f)>i/5>K(x_i)/6$.
\end{proof}
Note that we favoured simplicity of the proof over the strength of the bounds: the constants can be improved.

@@
